---
title: "Add a title page to your workflow report"
resource_files:
  - figures
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Add a title page to your workflow report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
source(system.file("extdata", "vignette-helpers.R", package = "ospsuite.reportingengine"))
```

```{r setup}
require(ospsuite.reportingengine)
```

In Mean Model and Population workflows, title pages can be included using the __`reportTitle`__ input argument.

The example below will be used to illustrate the available options when adding a title page to the final report.

<details>
<summary> __Code__ </summary>

```{r mini model example}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "Concentration of A",
  displayUnit = "nmol/ml"
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)

```

</details>

## Add a title only

When only a title is needed as a title page, the workflow will internally add the markdown title tag, __`"#"`__, to the __`reportTitle`__ as illustrated below.

<details>
<summary> __Code__ </summary>

```{r workflow title only}
# Create the workflow instance with the report title
workflow <-
  MeanModelWorkflow$new(
    simulationSets = setA,
    workflowFolder = "Example-1",
    reportTitle = "A meaningful title"
  )

# Set the workflow tasks to be run
workflow$activateTasks(c("simulate", "plotTimeProfilesAndResiduals"))

# Run the workflow
workflow$runWorkflow()
```

</details>

```{r example title only report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow))
```

## Add a title page

When the length of `reportTitle` is longer than 1, the workflow will assume `reportTitle` is a more advanced title page already formatted for markdown.
In such cases, `reportTitle` will used as is.

In the example below, the content of a more advanced title page is defined.
The corresponding page includes 

- A reference anchor that could be linked using `[Title page](#title-page)`
- A title with a markdown title tag
- A table (using `kable` for markdown formatting)

<details>
<summary> __Code__ </summary>

```{r title page content}

titlePage <- c(
  anchor("title-page"),
  "",
  "# A meaningful title",
  "",
  knitr::kable(
    data.frame(
      Date = Sys.Date(),
      Author = "OSP Suite - Reporting Engine",
      Subject = "Example Workflow"
    )
  )
)

```

```{r workflow advanced title}
# Here, it is more optimal to re-use the previous workflow 
# since only the report title page is changed and the same results are used
workflow$inactivateTasks("simulate")

workflow$reportTitle <- titlePage

# Re-run the workflow with the new title page
workflow$runWorkflow()
```

</details>

```{r example advanced title report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow))
```

## Use a file as title page

Another option is to use a markdown file as cover page.
In this case, the file path can directly be defined in `reportTitle` and the workflow will internally check that the file exists and include its content.

The example below save the previously defined title page as a file named :

<details>
<summary> __Code__ </summary>

```{r save title page}
titlePageFile <- "title-page.md"
write(
  x = titlePage,
  file = titlePageFile
)
```

```{r workflow title file}
# Here, it is more optimal to re-use the previous workflow 
# since only the report title page is changed and the same results are used
workflow$reportTitle <- titlePageFile

# Re-run the workflow with the new title page
workflow$runWorkflow()
```

</details>

```{r example title file report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow))
```

<details>
<summary> __Note__ </summary>
Note that running the workflow won't delete the title page.
The title file can be re-used if the workflow needs to be re-run.

```{r title file not deleted}
file.exists(titlePageFile)
```

```{r delete example folder, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow$workflowFolder, recursive = TRUE)
unlink(titlePageFile, recursive = TRUE)
```

</details>