---
title: "Demography Parameters in Population Workflows"
resource_files:
  - figures
output: 
  rmarkdown::html_vignette:
   number_sections: true
   toc: true
vignette: >
  %\VignetteIndexEntry{Demography Parameters in Population Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
source(system.file("extdata", "vignette-helpers.R", package = "ospsuite.reportingengine"))
```

```{r setup}
library(ospsuite.reportingengine)
```


# Overview

In Population workflows, the Demography task (`plotDemography`) aims at reporting the distributions of requested Demography parameters that will account for the population workflow type (as defined by variable `PopulationWorkflowTypes`).
The distributions are reported graphically using histograms or range plots depending on the workflow settings.

# Inclusion of observed data

Observed data can be included in the demography plots for comparison.

- In [Parallel and Ratio Comparison histograms](#histo-parallel), observed demography distributions will be displayed along with the simulated demography distribution.
- In [Pediatric histograms](#histo-pediatric), the distributions of all the simulated populations are displayed together in the same figure, observed distributions of all the populations are also displayed together in the same figure.
- In [range plots](#range-parallel), observed demography distributions will be displayed along with the simulated demography distribution.

As illustrated below, `DataSource` objects are necessary to include observed data in demography plots.

<details>
<summary> __Code__ </summary>

```{r definition of observed data}
# Define file paths for pkml simulation and populations
dictionary <- system.file("extdata", "popDictionary.csv",
  package = "ospsuite.reportingengine"
)
dataFile <- system.file("extdata", "obsPop.csv",
  package = "ospsuite.reportingengine"
)

dataSource <- DataSource$new(
  dataFile = dataFile, 
  metaDataFile = dictionary,
  caption = "Demography vignette example"
  )

```

</details>

In this process, the dictionary input as `metaDataFile` is essential.
Three columns of the dictionary are used by the workflow in order to map and convert the demography parameters.

- **pathID**: Defines the path of the demography parameter to include
- **datasetColumn**: Defines the column name of the demography parameter in your dataset
- **datasetUnit**: Defines the unit of the demography parameter in your dataset

The 2 next table shows the dictionary and the first rows of the observed data used in our example.

<details>
<summary> __Dictionary__ </summary>

```{r, echo=FALSE, results='asis'}
knitr::kable(readObservedDataFile(dictionary))
```

</details>

<details>
<summary> __Observed Data__ </summary>

```{r, echo=FALSE, results='asis'}
knitr::kable(head(readObservedDataFile(dataFile)))
```

</details>

**Tip**: to ensure that the unit defined in **datasetUnit** is appropriate, you can leverage the `{ospsuite}` package and copy/paste the value from R. Below provides a sample to get the appropriate unit for BMI (you need to use `format = 13` in order to keep correct encoded text as done for the square below).

```r
writeClipboard(ospsuite::ospUnits$BMI$`kg/m²`, format = 13)
```

# Vignette example

The code below creates the simulation sets used by the workflow defining the simulation, population and data files that will be used through this vignette.

In this example, the same observed data are included using the `dataSource` argument and different data are selected using the `dataSelection` argument of the  [`PopulationSimulationSet`](../reference/PopulationSimulationSet.html).


<details>
<summary> __Code__ </summary>

```{r definition of global variables}
# Define file paths for pkml simulation and populations
simulationFile <- system.file("extdata", "Aciclovir.pkml", package = "ospsuite")
adultFile <- system.file("extdata", "adults.csv", package = "ospsuite.reportingengine")
childrenFile <- system.file("extdata", "children.csv", package = "ospsuite.reportingengine")

adultSet <- PopulationSimulationSet$new(
  referencePopulation = TRUE,
  simulationSetName = "Adults",
  simulationFile = simulationFile,
  populationFile = adultFile,
  dataSource = dataSource,
  dataSelection = 'POP %in% "Adults"'
)

childrenSet <- PopulationSimulationSet$new(
  simulationSetName = "Children",
  simulationFile = simulationFile,
  populationFile = childrenFile,
  dataSource = dataSource,
  dataSelection = 'POP %in% "Children"'
)

```

</details>


# Parallel and Ratio Comparison Workflows

This section describes how to proceed with Parallel and Ratio Comparison workflows.

When creating a Parallel or a Ratio Comparison workflow, the `plotDemography` task includes 2 fields named `xParameters` and `yParameters`. 

- `xParameters` defines the demography parameters displayed in the x-axis of range plots.<br>
If no `xParameters` is defined (`xParameters = NULL`), histograms of demography parameters defined in `yParameters` are performed. 
- `yParameters` defines the demography parameters that will be displayed in the histograms if there is no `xParameters` or in the y-axis of range plots if `xParameters` defines demography parameters.

Note that the parameters can be defined as a vector of character values but also as a **named list in which case the name will be used as display name in the figures**.

By default, `xParameters = NULL` meaning that histogram are displayed. 
Users can check the default `xParameters` of their workflow type using the function [`getDefaultDemographyXParameters()`](../reference/getDefaultDemographyXParameters.html).

Default `yParameters` are independent of the workflow type and listed in the table below.

```{r, echo=FALSE, results='asis'}
knitr::kable(data.frame(
  "Parameter Path" = as.character(ospsuite.reportingengine:::DemographyDefaultParameters),
  "Display Name" = names(ospsuite.reportingengine:::DemographyDefaultParameters),
  check.names = FALSE
))
```

To define different demography parameters in `xParameters` and `yParameters`, the functions [`setXParametersForDemogrpahyPlot()`](../reference/setXParametersForDemogrpahyPlot.html) and [`setYParametersForDemogrpahyPlot()`](../reference/setYParametersForDemogrpahyPlot.html) can be respectively used.

## Histograms <a id="histo-parallel"></a>

The code below initializes the parallel comparison workflow and activate only the demography task.

<details>
<summary> __Code__ </summary>

```{r}
parallelWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$parallelComparison,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Parallel-Histograms",
  createWordReport = FALSE
)

inactivateWorkflowTasks(parallelWorkflow)
activateWorkflowTasks(parallelWorkflow, tasks = AllAvailableTasks$plotDemography)
```

</details>

To plot histograms of demography parameters, no demography parameters are defined in `xParameters` as illustrated in the example below which defines the parameters to be plotted in the histograms.

To define parameters not displayed by displayed, either a character array or a named list can be provided.
In the case of a named list, the name will be displayed while using its argument to query the appropriate demography parameter.
In this example, the parameter `Organism|Age` is defined with `Age` as displayed name.
User can also leverage the `{ospsuite}` package and look for the parameters in the enum `StandardPath`.

<details>
<summary> __Code__ </summary>

```{r}
# Query parameters defined in xParameters
getXParametersForDemogrpahyPlot(workflow = parallelWorkflow)
# Query parameters defined in yParameters
getYParametersForDemogrpahyPlot(workflow = parallelWorkflow)
```

```{r}
displayedParameters  <- list(
  Age = "Organism|Age", 
  Weight = "Organism|Weight", 
  Gender = "Gender"
  )

# Define no parameters in xParameters to get histograms
setXParametersForDemogrpahyPlot(workflow = parallelWorkflow, parameters = NULL)
# Define parameters to be displayed in histograms as yParameters
setYParametersForDemogrpahyPlot(workflow = parallelWorkflow, parameters = displayedParameters)
```

Run the corresponding workflow:

```{r}
parallelWorkflow$runWorkflow()
```

</details>

For Parallel and Ratio Comparison workflows, the demography histograms display the distributions for each simulation set along with its observed data if available ([#17](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/17), [#535](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/535)).

For categorical parameters, it can be noted that R factor levels are leveraged to display the name of the categories as is. 
Thus, users need to ensure that the same names are used in both their population files and observed data.

```{r example parallel histo report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(parallelWorkflow))
```

```{r delete parallel histo folder, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(parallelWorkflow$workflowFolder, recursive = TRUE)
```

## Range plots <a id="range-parallel"></a>

The code below initializes the parallel comparison workflow and activate only the demography task.

<details>
<summary> __Code__ </summary>

```{r}
parallelWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$parallelComparison,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Parallel-Range-Plots",
  createWordReport = FALSE
)

inactivateWorkflowTasks(parallelWorkflow)
activateWorkflowTasks(parallelWorkflow, tasks = AllAvailableTasks$plotDemography)
```

</details>

To display range plots of demography parameters, the demography parameters to display are defined in `xParameters` as illustrated in the example.
Note that **categorical** parameters defined as `xParameters` (_Gender_ in this case) are displayed ([#1088](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/1088) using box-whisker plots.

<details>
<summary> __Code__ </summary>

```{r}
xParameters  <- list(
  Age = "Organism|Age", 
  Gender = "Gender"
  )

setXParametersForDemogrpahyPlot(workflow = parallelWorkflow, parameters = xParameters)
setYParametersForDemogrpahyPlot(workflow = parallelWorkflow, parameters = displayedParameters)
```

Run the corresponding workflow:

```{r}
parallelWorkflow$runWorkflow()
```

</details>


For Parallel and Ratio Comparison workflows, the demography range plots display the distributions for each simulation set along with its observed data if available ([#17](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/17), [#535](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/535)).

Likewise, categorical parameters (*Gender* in this case) are displayed along with its observed data if available using boxplots instead ([#1088](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/1088).

For the comparison between simulated and observed populations to be the most relevant possible, the same binning is applied between simulated and observed populations.

```{r example parallel range report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(parallelWorkflow))
```

```{r delete parallel range folder, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(parallelWorkflow$workflowFolder, recursive = TRUE)
```

# Pediatric Workflows

This section describes how to proceed with Pediatric workflows.

When creating a Pediatric workflow, the `plotDemography` task includes 2 fields named `xParameters` and `yParameters`. 

- `xParameters` defines the demography parameters displayed in the x-axis of range plots.<br>
If no `xParameters` is defined (`xParameters = NULL`), histograms of demography parameters defined in `yParameters` are performed. 
- `yParameters` defines the demography parameters that will be displayed in the histograms if there is no `xParameters` or in the y-axis of range plots if `xParameters` defines demography parameters.

Note that the parameters can be defined as a vector of character values but also as a **named list in which case the name will be used as display name in the figures**.

By default, `xParameters = "Organism|Age"` meaning that range plots of demography parameters vs Age are displayed. 
Users can check the default `xParameters` of their workflow type using the function [`getDefaultDemographyXParameters()`](../reference/getDefaultDemographyXParameters.html).

Default `yParameters` are independent of the workflow type and listed in the table below.

```{r, echo=FALSE, results='asis'}
knitr::kable(data.frame(
  "Parameter Path" = as.character(ospsuite.reportingengine:::DemographyDefaultParameters),
  "Display Name" = names(ospsuite.reportingengine:::DemographyDefaultParameters),
  check.names = FALSE
))
```

To define different demography parameters in `xParameters` and `yParameters`, the functions [`setXParametersForDemogrpahyPlot()`](../reference/setXParametersForDemogrpahyPlot.html) and [`setYParametersForDemogrpahyPlot()`](../reference/setYParametersForDemogrpahyPlot.html) can be respectively used.

## Histograms <a id="histo-pediatric"></a>

The code below initializes the pediatric workflow and activate only the demography task.

<details>
<summary> __Code__ </summary>

```{r}
pediatricWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$pediatric,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Pediatric-Histograms",
  createWordReport = FALSE
)

inactivateWorkflowTasks(pediatricWorkflow)
activateWorkflowTasks(pediatricWorkflow, tasks = AllAvailableTasks$plotDemography)
```

</details>

To plot histograms of demography parameters, no demography parameters are defined in `xParameters` as illustrated in the example below which defines the parameters to be plotted in the histograms.

<details>
<summary> __Code__ </summary>

```{r}
setXParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = NULL)
setYParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = displayedParameters)
```

Run the corresponding workflow:

```{r}
pediatricWorkflow$runWorkflow()
```

</details>

For Pediatric workflows, the demography histograms display the distributions for all simulation sets together  ([#17](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/17), [#535](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/535)).
If observed data is available, the distributions for all simulation sets observed data are also displayed together.

For categorical parameters, it can be noted that R factor levels are leveraged to display the name of the categories as is. 
Thus, users need to ensure that the same names are used in both their population files and observed data.

```{r example pediatric histo report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(pediatricWorkflow))
```

```{r delete pediatric histo folder, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(pediatricWorkflow$workflowFolder, recursive = TRUE)
```


## Range plots

The code below initializes the pediatric workflow and activate only the demography task.

<details>
<summary> __Code__ </summary>

```{r}
pediatricWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$pediatric,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Pediatric-Range-Plots",
  createWordReport = FALSE
)

inactivateWorkflowTasks(pediatricWorkflow)
activateWorkflowTasks(pediatricWorkflow, tasks = AllAvailableTasks$plotDemography)
```

</details>

To display range plots of demography parameters, the demography parameters to display are defined in `xParameters` as illustrated in the example.
Note that **categorical** parameters defined as `xParameters` (_Gender_ in this case) are displayed ([#1088](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/1088) using box-whisker plots.

<details>
<summary> __Code__ </summary>

```{r}
xParameters  <- list(
  Age = "Organism|Age", 
  Gender = "Gender"
  )

setXParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = xParameters)
setYParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = displayedParameters)
```

Run the corresponding workflow:

```{r}
pediatricWorkflow$runWorkflow()
```

</details>


For Pediatric workflows, the demography range plots display the distributions for each simulation set along with its observed data if available ([#17](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/17), [#535](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/535)).
Additionally, range plots comparing the distribution of each simulation set to the **reference** simulation set is also performed.

Categorical parameters (*Gender* in this case) are also displayed along with the **reference** simulation set.

For the comparison between simulated and observed populations to be the most relevant possible, the same binning is used between simulated and observed populations.

For the additional comparison with the **reference** simulation set, the same binning is applied by default.
However, it is possible to display the global summary of the **reference** as horizontal ranges using the settings [**`referenceGlobalRange`**](#ref-global-range).


```{r example pediatric range report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(pediatricWorkflow))
```

```{r delete pediatric range folder, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(pediatricWorkflow$workflowFolder, recursive = TRUE)
```


# Demography settings

The demography plot task includes a few settings that may allow users to fine tune their demography plots.
Such settings are available in the field `workflow$plotDemography$settings`.

Since categorical parameters are displayed using boxplots, they are not affected the settings described below.

## Binning histograms and range plots

Currently, the default number of bins is set to **`r ospsuite.reportingengine:::AggregationConfiguration$bins`**.
However, users can update the number of bins or even include the bin edges within the field `bins` as illustrated below.

```r
workflow$plotDemography$settings$bins <- 5
```

In the example below, the binning is included within a parallel workflow:

<details>
<summary> __Code__ </summary>

```{r}
parallelWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$parallelComparison,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Parallel-Histograms",
  createWordReport = FALSE
)

inactivateWorkflowTasks(parallelWorkflow)
activateWorkflowTasks(parallelWorkflow, tasks = AllAvailableTasks$plotDemography)

# Only display histograms for Age, Weight and Gender
setYParametersForDemogrpahyPlot(workflow = parallelWorkflow, parameters = displayedParameters)

# Set up 5 bins for all histograms
parallelWorkflow$plotDemography$settings$bins <- 5

parallelWorkflow$runWorkflow()

```

</details>

In the report displayed below, the histograms are displayed with 5 bins except for categorical parameter whose categories are displayed as is.

```{r example parallel histogram report with bins, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(parallelWorkflow))
```

```{r delete parallel histogram report with bins, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(parallelWorkflow$workflowFolder, recursive = TRUE)
```


The same binning can also be applied to range plots as illustrated below:

<details>
<summary> __Code__ </summary>

```{r}
pediatricWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$pediatric,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Pediatric-Range",
  createWordReport = FALSE
)

inactivateWorkflowTasks(pediatricWorkflow)
activateWorkflowTasks(pediatricWorkflow, tasks = AllAvailableTasks$plotDemography)

# Only display range plots for Age and Weight (Gender being a categorical variable, it won't be displayed)
setYParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = displayedParameters)

# Set up 5 bins for all the range plots
pediatricWorkflow$plotDemography$settings$bins <- 5

pediatricWorkflow$runWorkflow()

```

</details>


```{r example pediatric range report with bins, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(pediatricWorkflow))
```

```{r delete pediatric range folder with bins, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(pediatricWorkflow$workflowFolder, recursive = TRUE)
```


Note that the algorithm that calculates the bin edges based on the number of bins aims at including the same number of points in each bin.

It is also possible to update default aggregation settings in order to update binning but also computed and displayed aggregation values using the following functions:

- [`setDefaultAggregationBins()`](../reference/setDefaultAggregationBins.html)
- [`setDefaultAggregationFunctions()`](../reference/setDefaultAggregationFunctions.html)
- [`setDefaultAggregationLabels()`](../reference/setDefaultAggregationLabels.html)

## Continuous vs stair step range plots

By default, demography range plots are displayed as stair step range plots.
It is however possible to connect the aggregated values to get a continuous range plot by setting the `stairstep` field to `FALSE` as shown below.

```r
workflow$plotDemography$settings$stairstep <- FALSE
```

In the example below, continuous range plots are included within a pediatric workflow:

<details>
<summary> __Code__ </summary>

```{r}
pediatricWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$pediatric,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Pediatric-Range",
  createWordReport = FALSE
)

inactivateWorkflowTasks(pediatricWorkflow)
activateWorkflowTasks(pediatricWorkflow, tasks = AllAvailableTasks$plotDemography)

# Only display range plots for Age and Weight (Gender being a categorical variable, it won't be displayed)
setYParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = displayedParameters)

# Set up continuous line between bins with stairstep settings being FALSE
pediatricWorkflow$plotDemography$settings$stairstep <- FALSE

pediatricWorkflow$runWorkflow()

```

</details>

```{r example pediatric range report stairstep, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(pediatricWorkflow))
```

```{r delete pediatric range folder stairstep, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(pediatricWorkflow$workflowFolder, recursive = TRUE)
```

## Dodging histogram bars

By default, demography histogram plots dodge the bars to prevent bars masking one another.
It is however possible to remove this settings in case the bars do not overlap by setting the `dodge` field to `FALSE` as shown below.

```r
workflow$plotDemography$settings$dodge <- FALSE
```

## Display reference population global range <a id="ref-global-range"></a>

Pediatric workflow range plots compare simulation sets to the **reference** simulation set ([#17](https://github.com/Open-Systems-Pharmacology/OSPSuite.ReportingEngine/issues/17)).

In cases where distributions in `xParameters` between populations are not comparable, it may be more appropriate to display only the global distributions of the `yParameters` of the reference population as horizontal lines/ranges.

As a consequence, it is possible to define such feature by setting the `referenceGlobalRange` field to `TRUE` as illustrated below.


```r
workflow$plotDemography$settings$referenceGlobalRange <- TRUE
```

In the example below, reference simulation set is displayed as its global range within a pediatric workflow:

<details>
<summary> __Code__ </summary>

```{r}
pediatricWorkflow <- PopulationWorkflow$new(
  workflowType = PopulationWorkflowTypes$pediatric,
  simulationSets = c(adultSet, childrenSet),
  workflowFolder = "Pediatric-Range",
  createWordReport = FALSE
)

inactivateWorkflowTasks(pediatricWorkflow)
activateWorkflowTasks(pediatricWorkflow, tasks = AllAvailableTasks$plotDemography)

# Only display range plots for Age and Weight (Gender being a categorical variable, it won't be displayed)
setYParametersForDemogrpahyPlot(workflow = pediatricWorkflow, parameters = displayedParameters)

# Set up reference global range, 
# only overall median 5th-95th percentiles of reference simulation set will be displayed
pediatricWorkflow$plotDemography$settings$referenceGlobalRange <- TRUE

pediatricWorkflow$runWorkflow()

```

</details>

```{r example pediatric range report referenceGlobalRange, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(pediatricWorkflow))
```

```{r delete pediatric range folder referenceGlobalRange, include = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(pediatricWorkflow$workflowFolder, recursive = TRUE)
```

