---
title: "PK parameters"
output: 
  rmarkdown::html_vignette:
   number_sections: true
   toc: true
#   self_contained: false
#   keep_md: true
vignette: >
  %\VignetteIndexEntry{pk-parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r report output style, child="reportOutputStyle.css", eval=TRUE}
```

```{r vignette setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
require(ospsuite.reportingengine)
```

This vignette focuses on PK parameters plots and tables generated using either mean model or population workflows.

# Overview

The PK parameter task aims at generating tables and, when possible, plots of requested PK parameters.
Depending on the workflow type, different plots and tables can be obtained.

## Inputs

Results obtained from the `calculatePKParameters` task and stored as csv files within the subdirectory __PKAnalysisResults__ from the `workflowFolder` directory are required in order to perform the PK parameters (`plotPKParameters`) task.
As a consequence, if the workflow output folder does not include the appropriate simulations, the task `calculatePKParameters` needs to be __active__.

The objects `SimulationSet` (or `PopulationSimulationSet` for population workflows) and `Output` define which and how the simulated PK parameters will be reported.

## Outputs

### Mean model workflow

For mean model workflows, the PK parameters task generates directly the table of PK parameters requested in the requested `Output` field `pkParameters`.

### Population model workflow

The distribution of the calculated PK parameters for each requested output are plotted as Box Whisker plots along with a table.
By default, the 5th, 25th, 50th, 75th and 95th percentiles are plotted.

Different types of a population workflow will lead to different output:

- __Parallel Comparison__: PK parameters are plotted parallel in the Box Whisker plots without any __reference population__.
- __Ratio Comparison__: Same as Parallel Comparison with additional Box Whisker plots and tables of the ratio of the PK parameters to the reference population is calculated and plotted in Box Whisker plots.
- __Pediatric__: PK Parameters are also  plotted as range plots of the PK parameters vs. age and weight.

Regarding range, the task option __xParameters__ can be used to define which demography parameters to use in x-axis of range plots.
Parallel and ratio use a `NULL` value by default, however it is also possible to add range plots to a Parallel or ratio comparison workflow by updating the value of __xParameters__.

Regarding ratio comparison, the formula below is used to calculate the Box Whisker plot percentile values:
$$PK\ Parameter\ Percentile\ Ratio = \frac{PK\ Parameter\ Percentile\ in\ Population}{PK\ Parameter\ Percentile\ in\ Reference\ Population}$$

# Template examples

The following sections will introduce template scripts including PK parameter tasks as well as their associated reports.
Table 1 shows the features tested in each template script.

```{r table 1, echo = FALSE, results='asis'}
examplesTable <- data.frame(
  Example = seq(1, 9),
  Workflow = c(rep("Mean model", 4), rep("Population model", 5)),
  `Workflow Type` = c(rep("-", 4), c("Parallel", "Parallel", "Ratio", "Pediatric", "Ratio")),
  `Simulation Sets` = c(1, 1, 2, 2, 1, 2, 2, 2, 2),
  check.names = FALSE
)
knitr::kable(examplesTable, caption = "Table 1: Features tested by each template script")
```

## Example 1

Example 1 shows a basic example of PK parameters performed by a mean model workflow.
In this example, '__C_max__' and '__AUC_tEnd__' are calculated.
Since the task PK parameters requires results from the folder __PKAnalysisResults__, 
the tasks `simulate` and `calculatePKParameters` need to be performed.
The task `simulate` needs to be performed because the task `calculatePKParameters` is using the time profile simulation results.

```{r example 1}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)

# Create the workflow instance
workflow1 <-
  MeanModelWorkflow$new(
    simulationSets = setA,
    workflowFolder = "Example-1"
  )

# Set the workflow tasks to be run
workflow1$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow1$runWorkflow()
```

```{r example 1 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow1$reportFileName)
```

The output report for Example 1 is shown below. 
The table indicates the requested parameters in their base unit.

<div class = "reportOutput">
```{r example 1 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow1$reportFileName))
```
</div>

```{r delete example 1, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow1$workflowFolder, recursive = TRUE)
```


## Example 2

Example 2 shows how to perform workflows with user defined PK parameters.
In this example, '__MyAUC__' is added to the list of PK parameters.
To create user defined PK parameter, the function __addUserDefinedPKParameter()__ from the `ospsuite` package needs to be used.

```{r example 2}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

myAUC <- addUserDefinedPKParameter(name = "MyAUC", standardPKParameter = StandardPKParameter$AUC_tEnd)
myAUC$startTime <- 50
myAUC$endTime <- 80

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd", "MyAUC")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)

# Create the workflow instance
workflow2 <-
  MeanModelWorkflow$new(
    simulationSets = setA,
    workflowFolder = "Example-2"
  )

# Set the workflow tasks to be run
workflow2$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow2$runWorkflow()
```

```{r example 2 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow2$reportFileName)
```

The output report for Example 2 is shown below. 
The table indicates the requested parameters in their base unit.

<div class = "reportOutput">
```{r example 2 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow2$reportFileName))
```
</div>

```{r delete example 2, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow2$workflowFolder, recursive = TRUE)
```

## Example 3

Example 3 shows a similar example with 2 simulation sets.
In this example, the names and units of the PK parameters are updated using the function __updatePKParameter__ from the `ospsuite` package.

```{r example 3}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax",
                  displayUnit = "nmol/l")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC",
                  displayUnit = "nmol*min/l")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)
outputB <- Output$new(
  path = "Organism|B|Concentration in container",
  displayName = "B",
  pkParameters = c("C_max", "AUC_tEnd")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)
setB <- SimulationSet$new(
  simulationSetName = "B",
  simulationFile = simulationFile,
  outputs = outputB
)

# Create the workflow instance
workflow3 <-
  MeanModelWorkflow$new(
    simulationSets = c(setA, setB),
    workflowFolder = "Example-3"
  )

# Set the workflow tasks to be run
workflow3$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow3$runWorkflow()
```

```{r example 3 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow3$reportFileName)
```

The output report for Example 3 is shown below. 
The table indicates the requested parameters in the updated units and display names.

<div class = "reportOutput">
```{r example 3 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow3$reportFileName))
```
</div>

```{r reset pk parameters from example 3, include=FALSE, echo=FALSE}
# Go back to previous PK parameter settings
updatePKParameter("C_max", 
                  displayName = "C_max",
                  displayUnit = "µmol/l")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC_tEnd",
                  displayUnit = "µmol*min/l")
```

## Example 4

Example 4 shows a different way of updating the PK parameters using the instance __PkParameterInfo__.
Using that instance, it is possible to define different names and units for the same parameters from different paths or simulations.

```{r example 4}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

CmaxA <- PkParameterInfo$new("C_max", 
                  displayName = "Cmax from A",
                  displayUnit = "µmol/l")
CmaxB <- PkParameterInfo$new("C_max", 
                  displayName = "Cmax from B",
                  displayUnit = "nmol/l")
AUCA <- PkParameterInfo$new("AUC_tEnd", 
                  displayName = "AUC from A",
                  displayUnit = "µmol*min/l")
AUCB <- PkParameterInfo$new("AUC_tEnd", 
                  displayName = "AUC from B",
                  displayUnit = "nmol*min/l")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c(CmaxA, AUCA)
)
outputB <- Output$new(
  path = "Organism|B|Concentration in container",
  displayName = "B",
  pkParameters = c(CmaxB, AUCB)
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)
setB <- SimulationSet$new(
  simulationSetName = "B",
  simulationFile = simulationFile,
  outputs = outputB
)

# Create the workflow instance
workflow4 <-
  MeanModelWorkflow$new(
    simulationSets = c(setA, setB),
    workflowFolder = "Example-3"
  )

# Set the workflow tasks to be run
workflow4$activateTasks("plotPKParameters")
workflow4$inactivateTasks(c("simulate", "calculatePKParameters"))

# Run the workflow
workflow4$runWorkflow()
```

```{r example 4 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow4$reportFileName)
```

The output report for Example 4 is shown below. 
The table indicates the requested parameters in the updated units and display names.

<div class = "reportOutput">
```{r example 4 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow4$reportFileName))
```
</div>

```{r delete example 4, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow4$workflowFolder, recursive = TRUE)
```

## Example 5

Example 5 shows pk parameters for a population workflow.
In this example, only one population-simulation pair is used.
Consequently, the workflow type needs to be a parallel comparison (no reference population).


```{r example 5}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)
populationFile <- system.file("extdata", "Pop500_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

set500 <- PopulationSimulationSet$new(
  simulationSetName = "Population A",
  simulationFile = simulationFile,
  populationFile = populationFile,
  outputs = outputA
)

# Create the workflow instance
workflow5 <-
  PopulationWorkflow$new(
    workflowType = PopulationWorkflowTypes$parallelComparison,
    simulationSets = set500,
    workflowFolder = "Example-5"
  )

# Set the workflow tasks to be run
workflow5$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow5$runWorkflow()
```

```{r example 5 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow5$reportFileName)
```

The output report for Example 6 is shown below. 
For the population, Box Whisker plots in linear and log scales are.

<div class = "reportOutput">
```{r example 5 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow5$reportFileName))
```
</div>

```{r delete example 5, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow5$workflowFolder, recursive = TRUE)
```

## Example 6

Example 6 shows pk parameters for a population workflow.
In this example, 2 population-simulation pairs are now used.
Consequently, the workflow type needs to be a parallel comparison (no reference population).

```{r example 6}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)
populationFile500 <- system.file("extdata", "Pop500_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)
populationFile200 <- system.file("extdata", "Pop200_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

set500 <- PopulationSimulationSet$new(
  simulationSetName = "500 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile500,
  outputs = outputA
)
set200 <- PopulationSimulationSet$new(
  simulationSetName = "200 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile200,
  outputs = outputA
)

# Create the workflow instance
workflow6 <-
  PopulationWorkflow$new(
    workflowType = PopulationWorkflowTypes$parallelComparison,
    simulationSets = c(set500, set200),
    workflowFolder = "Example-6"
  )

# Set the workflow tasks to be run
workflow6$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow6$runWorkflow()
```

```{r example 6 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow6$reportFileName)
```

The output report for Example 6 is shown below. 
For the population, BoxWhisker plots in linear and log scales are.

<div class = "reportOutput">
```{r example 6 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow6$reportFileName))
```
</div>

## Example 7

Example 7 is the same as the previous example, only this time a ratio type population workflow is performed.
As a consequence, the workflow type was specified when creating the workflow instance.
In this workflow type, a __reference population__ is needed.
Since the workflows can re-use the results from the previous PK analysis, only the plot task needs to be re-run.

```{r example 7}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)
populationFile500 <- system.file("extdata", "Pop500_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)
populationFile200 <- system.file("extdata", "Pop200_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

set500 <- PopulationSimulationSet$new(
  referencePopulation = TRUE,
  simulationSetName = "500 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile500,
  outputs = outputA
)
set200 <- PopulationSimulationSet$new(
  simulationSetName = "200 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile200,
  outputs = outputA
)


# Create the workflow instance
workflow7 <-
  PopulationWorkflow$new(
    workflowType = PopulationWorkflowTypes$ratioComparison,
    simulationSets = c(set500, set200),
    workflowFolder = "Example-6"
  )

# Set the workflow tasks to be run
workflow7$inactivateTasks()
workflow7$activateTasks("plotPKParameters")

# Run the workflow
workflow7$runWorkflow()
```

```{r example 7 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow7$reportFileName)
```

The output report for Example 7 is shown below. 

<div class = "reportOutput">
```{r example 7 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow7$reportFileName))
```
</div>

## Example 8

Example 8 is the same as the 2 previous examples, this time a pediatric type population workflow is performed.
As a consequence, the workflow type was specified when creating the workflow instance.
In this workflow type, a __reference population__ is needed.
In the test population, age and body weight are not population parameters only *p1*, *p2* and *p3*, as a consequence, *p3* was used in the x-axis in the range plots.
These parameters will be used instead of age for the range plots.
Since the workflows can re-use the results from the previous PK analysis, only the plot task needs to be re-run.

```{r example 8}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)
populationFile500 <- system.file("extdata", "Pop500_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)
populationFile200 <- system.file("extdata", "Pop200_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

set500 <- PopulationSimulationSet$new(
  referencePopulation = TRUE,
  simulationSetName = "500 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile500,
  outputs = outputA
)
set200 <- PopulationSimulationSet$new(
  simulationSetName = "200 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile200,
  outputs = outputA
)


# Create the workflow instance
workflow8 <-
  PopulationWorkflow$new(
    workflowType = PopulationWorkflowTypes$pediatric,
    simulationSets = c(set500, set200),
    workflowFolder = "Example-6"
  )

# Set the workflow tasks to be run
workflow8$inactivateTasks()
workflow8$activateTasks("plotPKParameters")

setXParametersForPkParametersPlot(workflow8, "Organism|p3")

# Run the workflow
workflow8$runWorkflow()
```

```{r example 8 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow8$reportFileName)
```

The output report for Example 8 is shown below. 

<div class = "reportOutput">
```{r example 8 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow8$reportFileName))
```
</div>

```{r delete example 8, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow8$workflowFolder, recursive = TRUE)
```

## Example 9

In some cases, it may be necessary to compare different PK parameters from different simulations.
Example 9 shows how to define a group identifier in `PkParameterInfo` objects.
In this example, a user defined PK parameter named `"MyAUC"` is created using the function `addUserDefinedPKParameter`.
Then, 2 `PkParameterInfo` objects are created with the same value for their field `group` while their field `pkParameter` indicates a different PK parameter.
If `group` is left undefined, it will take the value defined in `pkParameter` input argument by default.

The workflow created using this method will compare *"AUC_tEnd"* from the first population with *"myAUC"* from the second population.

```{r example 9}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)
populationFile500 <- system.file("extdata", "Pop500_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)
populationFile200 <- system.file("extdata", "Pop200_p1p2p3.csv",
  package = "ospsuite.reportingengine"
)

# Create a user defifened PK parameter
myAUC <- addUserDefinedPKParameter(name = "MyAUC", standardPKParameter = StandardPKParameter$AUC_tEnd)
myAUC$startTime <- 50
myAUC$endTime <- 80

# Define PkParameterInfo objects sharing the same groups
AUC500 <- PkParameterInfo$new(
  pkParameter = "AUC_tEnd", 
  group = 1,
  displayName = "AUC",
  displayUnit = "µmol*min/l")
AUC200 <- PkParameterInfo$new(
  pkParameter = "MyAUC", 
  group = 1,
  displayName = "AUC",
  displayUnit = "µmol*min/l")

# Define the input parameters
output500 <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = AUC500
)
output200 <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = AUC200
)

set500 <- PopulationSimulationSet$new(
  referencePopulation = TRUE,
  simulationSetName = "500 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile500,
  outputs = output500
)
set200 <- PopulationSimulationSet$new(
  simulationSetName = "200 individuals",
  simulationFile = simulationFile,
  populationFile = populationFile200,
  outputs = output200
)

# Create the workflow instance
workflow9 <-
  PopulationWorkflow$new(
    workflowType = PopulationWorkflowTypes$ratioComparison,
    simulationSets = c(set500, set200),
    workflowFolder = "Example-9"
  )

# Set the workflow tasks to be run
workflow9$activateTasks(c("simulate", "calculatePKParameters", "plotPKParameters"))
workflow9$activateTasks("plotPKParameters")

# Run the workflow
workflow9$runWorkflow()
```

```{r example 9 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow9$reportFileName)
```

The output report for Example 9 is shown below. 

<div class = "reportOutput">
```{r example 9 report, echo=FALSE}
htmltools::includeHTML(sub(".md", ".html", workflow9$reportFileName))
```
</div>

```{r delete example 9, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow9$workflowFolder, recursive = TRUE)
```