---
title: "PK parameters"
resource_files:
  - figures
output: 
  rmarkdown::html_vignette:
   number_sections: true
   toc: true
vignette: >
  %\VignetteIndexEntry{PK parameters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
source(system.file("extdata", "vignette-helpers.R", package = "ospsuite.reportingengine"))
```

```{r setup}
require(ospsuite.reportingengine)
```

This vignette focuses on PK parameters plots and tables generated using either mean model or population workflows.

# Overview

The PK parameter task aims at generating tables and, when possible, plots of requested PK parameters.
Depending on the workflow type, different plots and tables can be obtained.

## Inputs

Results obtained from the `calculatePKParameters` task and stored as csv files within the subdirectory __PKAnalysisResults__ from the `workflowFolder` directory are required in order to perform the PK parameters (`plotPKParameters`) task.
As a consequence, if the workflow output folder does not include the appropriate simulations, the task `calculatePKParameters` needs to be __active__.

The objects `SimulationSet` (or `PopulationSimulationSet` for population workflows) and `Output` define which and how the simulated PK parameters will be reported.

## Outputs

### Mean model workflow

For mean model workflows, the PK parameters task generates directly the table of PK parameters requested in the requested `Output` field `pkParameters`.

### Population model workflow

The distribution of the calculated PK parameters for each requested output are plotted as Box Whisker plots along with a table.
By default, the 5th, 25th, 50th, 75th and 95th percentiles are plotted.

Check the article [PK Parameters in Population Workflows](pop-pk-parameters.html) for more details

# Template examples

The following sections will introduce template scripts including PK parameter tasks as well as their associated reports.
Table 1 shows the features tested in each template script.
All the examples use Mean Model Workflows.

```{r table 1, echo = FALSE, results='asis'}
examplesTable <- data.frame(
  Example = seq(1, 4),
  Workflow = "MeanModelWorkflow",
  `Simulation Sets` = c(1, 1, 2, 2),
  check.names = FALSE
)
knitr::kable(examplesTable, caption = "Table 1: Features tested by each template script")
```

## Example 1

Example 1 shows a basic example of PK parameters performed by a mean model workflow.
In this example, '__C_max__' and '__AUC_tEnd__' are calculated.
Since the task PK parameters requires results from the folder __PKAnalysisResults__, 
the tasks `simulate` and `calculatePKParameters` need to be performed.
The task `simulate` needs to be performed because the task `calculatePKParameters` is using the time profile simulation results.

<details>
<summary> __Code__ </summary>

```{r example 1}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)

# Create the workflow instance
workflow1 <-
  MeanModelWorkflow$new(
    simulationSets = setA,
    workflowFolder = "Example-1"
  )

# Set the workflow tasks to be run
workflow1$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow1$runWorkflow()
```

</details>

The output report for Example 1 is shown below. 
The table indicates the requested parameters in their base unit.

```{r example 1 report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow1))
```

```{r delete example 1, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow1$workflowFolder, recursive = TRUE)
```


## Example 2

Example 2 shows how to perform workflows with user defined PK parameters.
In this example, '__MyAUC__' is added to the list of PK parameters.
To create user defined PK parameter, the function __addUserDefinedPKParameter()__ from the `ospsuite` package needs to be used.

<details>
<summary> __Code__ </summary>

```{r example 2}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

myAUC <- addUserDefinedPKParameter(name = "MyAUC", standardPKParameter = StandardPKParameter$AUC_tEnd)
myAUC$startTime <- 50
myAUC$endTime <- 80

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd", "MyAUC")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)

# Create the workflow instance
workflow2 <-
  MeanModelWorkflow$new(
    simulationSets = setA,
    workflowFolder = "Example-2"
  )

# Set the workflow tasks to be run
workflow2$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow2$runWorkflow()
```

</details>

The output report for Example 2 is shown below. 
The table indicates the requested parameters in their base unit.

```{r example 2 report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow2))
```

```{r delete example 2, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow2$workflowFolder, recursive = TRUE)
```

## Example 3

Example 3 shows a similar example with 2 simulation sets.
In this example, the names and units of the PK parameters are updated using the function __updatePKParameter__ from the `ospsuite` package.

<details>
<summary> __Code__ </summary>

```{r example 3}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFileA <- system.file("extdata", "MiniModel1.pkml",
  package = "ospsuite.reportingengine"
)

simulationFileB <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

updatePKParameter("C_max", 
                  displayName = "Cmax",
                  displayUnit = "nmol/l")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC",
                  displayUnit = "nmol*min/l")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c("C_max", "AUC_tEnd")
)
outputB <- Output$new(
  path = "Organism|B|Concentration in container",
  displayName = "B",
  pkParameters = c("C_max", "AUC_tEnd")
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFileA,
  outputs = outputA
)
setB <- SimulationSet$new(
  simulationSetName = "B",
  simulationFile = simulationFileB,
  outputs = outputB
)

# Create the workflow instance
workflow3 <-
  MeanModelWorkflow$new(
    simulationSets = c(setA, setB),
    workflowFolder = "Example-3"
  )

# Set the workflow tasks to be run
workflow3$activateTasks(c("simulate", 
                          "calculatePKParameters", 
                          "plotPKParameters"))

# Run the workflow
workflow3$runWorkflow()
```

</details>

The output report for Example 3 is shown below. 
The table indicates the requested parameters in the updated units and display names.

```{r example 3 report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow3))
```

```{r reset pk parameters from example 3, include=FALSE, echo=FALSE}
# Go back to previous PK parameter settings
updatePKParameter("C_max", 
                  displayName = "C_max",
                  displayUnit = "µmol/l")
updatePKParameter("AUC_tEnd", 
                  displayName = "AUC_tEnd",
                  displayUnit = "µmol*min/l")
```

## Example 4

Example 4 shows a different way of updating the PK parameters using the instance __PkParameterInfo__.
Using that instance, it is possible to define different names and units for the same parameters from different paths or simulations.

<details>
<summary> __Code__ </summary>

```{r example 4}
# Get the pkml simulation file: "MiniModel2.pkml"
simulationFile <- system.file("extdata", "MiniModel2.pkml",
  package = "ospsuite.reportingengine"
)

CmaxA <- PkParameterInfo$new("C_max", 
                  displayName = "Cmax from A",
                  displayUnit = "µmol/l")
CmaxB <- PkParameterInfo$new("C_max", 
                  displayName = "Cmax from B",
                  displayUnit = "nmol/l")
AUCA <- PkParameterInfo$new("AUC_tEnd", 
                  displayName = "AUC from A",
                  displayUnit = "µmol*min/l")
AUCB <- PkParameterInfo$new("AUC_tEnd", 
                  displayName = "AUC from B",
                  displayUnit = "nmol*min/l")

# Define the input parameters
outputA <- Output$new(
  path = "Organism|A|Concentration in container",
  displayName = "A",
  pkParameters = c(CmaxA, AUCA)
)
outputB <- Output$new(
  path = "Organism|B|Concentration in container",
  displayName = "B",
  pkParameters = c(CmaxB, AUCB)
)

setA <- SimulationSet$new(
  simulationSetName = "A",
  simulationFile = simulationFile,
  outputs = outputA
)
setB <- SimulationSet$new(
  simulationSetName = "B",
  simulationFile = simulationFile,
  outputs = outputB
)

# Create the workflow instance
workflow4 <-
  MeanModelWorkflow$new(
    simulationSets = c(setA, setB),
    workflowFolder = "Example-3"
  )

# Set the workflow tasks to be run
workflow4$activateTasks("plotPKParameters")
workflow4$inactivateTasks(c("simulate", "calculatePKParameters"))

# Run the workflow
workflow4$runWorkflow()
```

</details>

The output report for Example 4 is shown below. 
The table indicates the requested parameters in the updated units and display names.

```{r example 4 report, echo=FALSE, results='asis'}
cat(includeReportFromWorkflow(workflow4))
```

```{r example 4 render, include=FALSE, echo=FALSE}
# Convert the markdown document into html
# Otherwise the root path is messed up and cannot print the report correctly
rmarkdown::render(workflow4$reportFilePath)
```

```{r delete example 4, inlcude = FALSE, echo=FALSE}
# Remove the workflow folders
unlink(workflow4$workflowFolder, recursive = TRUE)
```
