---
title: "Qualification Workflow"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Qualification Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
require(ospsuite.reportingengine)
```

# Objectives

This vignette aims at introducing qualification workflows.
Qualification workflows are usually build to derive Qualification Reports that are recreated with every new OSP Release.

# Qualification runner

The **QualificationRunner** is the software in charge of initializing and managing the qualification workflow by creating and organizing the inputs required for the reporting engine.

The function `startQualificationRunner`, inspired from its Matlab version, launches the **QualificationRunner** on a project.
The function requires the inputs below:

- `qualificationRunnerFolder`: the directory where the file **QualificationRunner.exe** is located
- `qualificationPlanFile`: the path of the input qualification plan (json file summarizing the project qualification)
- `outputFolder`: the path of the output directory created by the qualification runner

```{r qualification runner, eval = FALSE}
startQualificationRunner(qualificationRunnerFolder, qualificationPlanFile, outputFolder)
```

Note that, in the R version of `startQualificationRunner`, the simulations and PK analyses are not run from the **QualificationRunner** but from the workflow itself.
This update aims at preventing the runs of all the simulations available in the qualification.
Only the simulations results and PK analyses required by the configuration plan are performed to optimize the building time of a qualification report.

# Configuration Plan

The configuration plan is included in a json file, created by the qualification runner, that defines the configuration of the final report.
The function `loadConfigurationPlan` reads such a file and creates a `ConfigurationPlan` object that all the knowledge required to build the final report.
It is internally called upon creation of a `QualificationWorkflow` and requires the inputs below:

- `configurationPlanFile`: path of the Configuration Plan json file used by the Qualification workflow
- `workflowFolder`: path of the output directory created and/or used by the Workflow.

# Qualification workflow

Qualification workflows are built from `Qualificationworkflow` objects which derive from `Workflow` objects.
As such, they are also based on the concept of tasks but show a difference in the way they are created.

## How to create a qualification workflow

The function `loadQualificationWorkflow` is advised to create a `QualificationWorkflow` object from a configuration plan.
It requires the same inputs as `loadConfigurationPlan` (`workflowFolder` and `configurationPlanFile`).

```{r load qualification workflow, eval = FALSE}
workflow <- loadQualificationWorkflow(workflowFolder, configurationPlanFile)
```

## Qualification workflow features

### Configuration Plan

All `QualificationWorkflow` objects inlcude a `configurationPlan` field.
This field is a `ConfigurationPlan` object that provides a smart interface for buidling the report.

In particular, all the fields below are reorganized in a data.frame format clarifying the design of the configuration plan.

- `simulationMappings`
- `observedDataSets`
- `sections`

Additionally, the `configurationPlan` field includes many `get` methods to search for specific paths of files managedd by the configuration plan (`$getSimulationPath(project, simulation)`, `$getObservedDataPath(id)`, etc)

### Tasks available for Qualification workflow

Two simulation tasks are available and active by default in the qualification workflows:

- `simulate`: simulate time profile results requested in the report plots
- `calculatePKParameters`: calculate the PK Parameters requested in the PK and DDI ratio plots

Their results are stored within the `workflowFolder` output directory, in *SimulationResults* and *PKAnalysisResults* directories respectively.

Five plot tasks are available in the qualification workflows, they are only active if defined by the configuration plan:

- `plotTimeProfiles`: plot time profiles of the simulated outputs and/or observed data.
- `plotGOFMerged`: plot goodness of fit results of grouped simulated outputs vs observed data.
- `plotComparisonTimeProfile`: plot comparison of time profiles from the simulated outputs and/or observed data.
- `plotPKRatio`: plot PK ratios of simulated vs observed PK parameters as a function of covariate (usually Age).
- `plotDDIRatio`: plot Drug-Drug Interaction ratios of simulated vs observed PK parameters.

### Settings and artifacts

Most of the settings and artifacts are read from the configuration plan file and translated into task settings.
Settings for simulations include `showProgress` to show the progress of the `simulate` task and `allowedCores` to perform the simulations in parallel.

Plot tasks can include multiple results for the report, usually called artifacts.
If the field **Artifacts** is not mentioned in a plot of the configuration plan, all its corresponding artifacts results will be added to the report.

#### predictedVsObserved and residualsOverTime

For `plotGOFMerged` and `plotDDIRatio` tasks, the plots represent observed vs predicted data.
It is consequently possible to include them as predicted vs observed and/or residuals vs time/predicted by including the options **predictedVsObserved**, **residualsOverPredicted** and **residualsOverTime** in the field **PlotTypes** of the configuration plan.

#### Gemotric Mean Fold Error (GMFE)

For `plotGOFMerged`, `plotPKRatio` and `plotDDIRatio` tasks, it is possible to include a **GMFE** artifact in the configuration plan.
This artifact corresponds to the addition of a table showing the Gemotric Mean Fold Error (*GMFE*) of the *residuals* calculated for specific plot using the formula $GMFE = 10^{\Sigma log_{10}(residuals)}$.

#### Measure

For `plotPKRatio` and `plotDDIRatio` tasks, it is possible to include a **Measure** artifact in the configuration plan.
This artifact corresponds to the addition of a table showing:

- For `plotPKRatio`, count and percent of ratios lying within 1.5- and 2-fold error ranges.
- For `plotDDIRatio`, count and percent of ratios lying within 2-fold and Guest *et al.* error ranges.

#### Table

For `plotPKRatio` task, it is possible to include a **Table** artifact in the configuration plan.
This artifact corresponds to the addition of a table showing comprehensive observed and simulatied PK data of the PK Ratio analysis.

#### Settings

Since plot tasks can include numerical results, `digits` and `scientific` are fields available in each task `settings` that will define the format of their numeric values.

## How to update and re-run a qualification workflow

If simulations need to be re-run, it is advised to relaod the workflow using `loadQualificationWorkflow`.

In case the update does not require to re-run simulations (e.g. updating the section content, modifying plot aesthetics or removing a plot), the `QualificationWorkflow` method `$updateConfigurationPlan(configurationPlanFile)` can be used to avoid the process of reloading and checking all the simulations of the workflow.

Then, you can also `inactivate` the tasks corresponding to the re-run of the simulations before re-running the workflow.

```{r rerun workflow, eval = FALSE}
# Reload Configuration Plan
workflow$updateConfigurationPlan(
  configurationPlanFile = newConfigurationPlanFile
)

# Inactivate simulations already performed
workflow$simulate$inactivate()
workflow$calculatePKParameters$inactivate()

# Re-run and build updated report
workflow$runWorkflow()
```

## Template

A template R script for qualification workflows is available and can be found using the function `system.file("extdata", "qualification-workflow-template.R", package = "ospsuite.reportingengine")`.
The content of the file is defined below.

```{r template, echo=FALSE, comment=NA}
cat(
  readLines(
    system.file(
      "extdata", "qualification-workflow-template.R",
      package = "ospsuite.reportingengine"
    )
  ),
  sep = "\n"
)
```
