---
title: "Sensitivity Analysis"
output: 
  rmarkdown::html_vignette:
   number_sections: true
vignette: >
  %\VignetteIndexEntry{sensitivity-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
require(ospsuite.reportingengine)
```

# Overview

The sensitivity task aims at generating plots showing PK parameters sensitivity to variations of model input parameters.

## Inputs

Results obtained from the `meanModelSensitivityAnalysis` or `populationSensitivityAnalysis` task and stored as csv files within the subdirectory __SensitivityResults__ from the `workflowFolder` directory are required in order to perform the sensitivity (`plotSensitivity`) task.
As a consequence, if the workflow output folder does not include the appropriate simulations, the task `meanModelSensitivityAnalysis` or `populationSensitivityAnalysis` needs to be __active__.

The objects `SimulationSet` (or `PopulationSimulationSet` for population workflows) and `Output` define which and how the simulated PK parameters will be reported.

## Ouptuts

### Mean model workflow

### Population model workflow

# Template examples
