---
title: "Workflow notes"
author: "OSPSuiteR 2019"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{workflow-vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
--- 

```{r, results='asis', echo=FALSE} 
library(ospsuite)
library(tlf)
library(ospsuite.reportingengine)
#source("./R/get-enum.R")
```

# Simple simulation from MoBi

Load a simulation:
```{r, results='asis'}
pkmlFilePath <- "../data/simpleMobiEventSim.pkml"
sim <- loadSimulation(pkmlFilePath) 
```

Print possible paths to simulate:
Load a simulation:
```{r}
sim$outputSelections
```


If another path is required, it can be added using the `addOutputs`:
```{r}
pathEnumList <- getEnum(pkmlFilePath)
addOutputs(pathEnumList$Organism$blockA$mol1$Concentration$path,simulation = sim)
sim$outputSelections
```


Run a simulation:
```{r}
res <- runSimulation(sim)
```

Export results of a simulation:
```{r, results='asis'}
ospsuite::exportResultsToCSV( results = res , filePath = "../data/exp1.csv")
```

Extract the results from the simulation results:
```{r}
timeValues    <- res$timeValues
blockAResults <- res$getValuesByPath(path = res$allQuantityPaths[1] , individualIds = 0)
blockBResults <- res$getValuesByPath(path = res$allQuantityPaths[2] , individualIds = 0)
print(head(timeValues))
print(head(blockAResults))
print(head(blockBResults))
```

Can also get a list which has a tree structure for each output:
```{r}
allResults <- getOutputValues(simulationResults = res,
                              quantitiesOrPaths = res$allQuantityPaths[c(1,2)],
                              individualIds = 0)
```


# Simulation from PKSim population model

Load a PKSim simulation:
```{r}
pkmlFilePath <- "../data/individualPksimSim.pkml"
indSim <- loadSimulation(pkmlFilePath) 
```


Load a population from a population datafile:
```{r}
popDataFilePath <- "../data/popData.csv"
population <- loadPopulation(csvPopulationFile = popDataFilePath)
```


Run a population simulation:
```{r}
popRes <- runSimulation(simulation = indSim , population = population)
```


Get population results in TLF (dataframe) format:
```{r}
allResultsPopTLF <- getOutputValuesTLF(simulationResults = popRes,
                                       population = population,
                                       quantitiesOrPaths = popRes$allQuantityPaths[c(1,2)],
                                       individualIds = seq(0,5)) 
print(head(allResultsPopTLF$data))
```


```{r, fig.height=5, fig.width=7.5}
wfGrouping <- Grouping$new( group = "IndividualId" )
wfDataMapping <- TimeProfileDataMapping$new(x = "Time",
                                            y = popRes$allQuantityPaths[1],
                                            groupMapping = GroupMapping$new(color = wfGrouping))
wfTimeProfilePlot <- plotTimeProfile(data = allResultsPopTLF$data, metaData = allResultsPopTLF$metaData, dataMapping = wfDataMapping)
show(wfTimeProfilePlot)
```



